---
title: 'IUCN red lists'
author: "Hans Van Calster"
date: "1 september 2017"
output:
  word_document: default
  pdf_document: default
  html_document: default
---
```{r setup, include = FALSE}
library(knitr)
opts_chunk$set(echo = FALSE)
library(tidyverse)
library(readxl)
library(vegan)
library(INBOmisc)
library(RODBC)
```

## Set working directory and get the data
```{r, echo = TRUE, include = TRUE}
setwd("C:/Users/dirk_maes/Google Drive/Publicaties/inprep2017_TraitRedListingFlandersAnalyses/")
pathFig <- "Figuren/"
```

```{r, echo = TRUE, include = TRUE}
MDB <- odbcConnectAccess2007("RedListFlandersAnalyses_Work.accdb")
rl_all1 <- sqlQuery(MDB, "select * from [qryCCA]")
rl_all2 <- sqlQuery(MDB, "select * from [qryCCA2]")
odbcClose(MDB)
```

If species have 2 biotopes, copy the second biotope into the Biotope column
```{r, echo = TRUE}
rl_all <- rbind(rl_all1, rl_all2)
nrow(rl_all)
```

## Count the number of species per biotope and eliminate biotopes with less than 5 species
```{r, echo = TRUE}
table(rl_all$Biotope)
rl_all <- rl_all %>%
	filter(Biotope != "Marine")
```

## Prepare a table for the cca analysis
```{r, echo = TRUE}
rl_all_wide <- rl_all %>%
	select(TaxonomicGroup, SpeciesnameAsPublished, RLC, Biome, Biotope, Cuddliness, Lifespan, Mobility, NutrientLevel) %>%
	filter(RLC != "DD", RLC != "DD(PT)") %>%
	group_by(RLC, Biome, Biotope, Cuddliness, Lifespan, Mobility, NutrientLevel) %>%
	summarise(nspec = n()) %>%
	spread(key = RLC, value = nspec, fill = 0) %>%
	ungroup() %>%
	filter(complete.cases(.))
#write.table(rl_all_wide, file = "rl_all_wide.csv", sep = ";", row.names = FALSE)
```

## Do CCA and optimise the model using anova
```{r, include  = TRUE}
cca1_all <- cca(rl_all_wide %>% select(CR:VU) ~ Biome
								+ Biotope
								+ Cuddliness
								+ Lifespan
								+ Mobility
								+ NutrientLevel
								, data = rl_all_wide)
summary(cca1_all)
anova(cca1_all, by = "terms")
```

## Make biplot and save as jpg
```{r, echo = TRUE, include = TRUE}
pbiplot_all <- ggbiplot_vegan(cca1_all, site_geom = "blank", cn_geom = "text", base_sizes = c(5,8,5))
pbiplot_all
ggsave(pbiplot_all, filename = "CCA_biplot_All.jpg", dpi = 600, width = 12, height = 12)
```

# Do analysis for **invertebrates** only
```{r, echo = TRUE}
rl_invertebrates <- rl_all %>% 
	filter(Spine == "Invertebrate")
nrow(rl_invertebrates)

table(rl_invertebrates$Biotope)

rl_invertebrates <- rl_invertebrates %>% 
	filter(Biotope != "Agriculture" & Biotope != "Anthropogenic" & Biotope != "Marine")
nrow(rl_invertebrates)

table(rl_invertebrates$Biotope)
```

```{r, echo = TRUE}
rl_invertebrates_wide <- rl_invertebrates %>%
	select(TaxonomicGroup, SpeciesnameAsPublished, RLC, Biome, Biotope, Cuddliness, Lifespan, Mobility, NutrientLevel) %>%  
	filter(RLC != "DD", RLC != "DD(PT)") %>%
	group_by(RLC, Biome, Biotope, Cuddliness, Mobility, NutrientLevel, Lifespan) %>%
	summarise(nspec = n()) %>%
	spread(key = RLC, value = nspec, fill = 0) %>%
	ungroup() %>%
	filter(complete.cases(.))
```

Do not include Lifespan in the analysis, since all invertebrates are shortlived

```{r, include = TRUE}
cca1_invertebrates <- cca(rl_invertebrates_wide %>% select(CR:VU) ~ Biome
													+ Biotope
													+ Cuddliness
													#+ Lifespan
													+ Mobility
													+ NutrientLevel
													, data = rl_invertebrates_wide)
summary(cca1_invertebrates)
anova(cca1_invertebrates, by = "terms")
```

```{r, echo = TRUE}
pbiplot_invertebrates <- ggbiplot_vegan(cca1_invertebrates, site_geom = "blank", cn_geom = "text", base_sizes = c(5, 8, 5))
pbiplot_invertebrates
ggsave(pbiplot_invertebrates, filename = "CCA_biplot_Invertebrates.jpg", dpi = 600, width = 10, height = 10)
```

# Do analysis for **vertebrates** only
```{r, echo = TRUE}
rl_vertebrates <- rl_all %>% 
	filter(Spine == "Vertebrate")
table(rl_vertebrates$Biotope)
nrow(rl_vertebrates)
```

```{r, echo = TRUE}
rl_vertebrates_wide <- rl_vertebrates %>%
select(TaxonomicGroup, SpeciesnameAsPublished, RLC, Biome, Biotope, Cuddliness, Lifespan, Mobility, NutrientLevel) %>%
	filter(RLC != "DD", RLC != "DD(PT)") %>%
	group_by(RLC, Biome, Biotope, Lifespan, Cuddliness, Mobility, NutrientLevel) %>%
	summarise(nspec = n()) %>%
	spread(key = RLC, value = nspec, fill = 0) %>%
	ungroup() %>%
	filter(complete.cases(.))
```

```{r, include = TRUE}
cca1_vertebrates <- cca(rl_vertebrates_wide %>% select(CR:VU) ~ #Biome
												+ Biotope
												+ Cuddliness
												+ Lifespan
												+ Mobility
												+ NutrientLevel
												, data = rl_vertebrates_wide) 
summary(cca1_vertebrates)
anova(cca1_vertebrates, by = "terms")
```

```{r, echo = TRUE}
pbiplot_vertebrates <- ggbiplot_vegan(cca1_vertebrates, site_geom = "blank", cn_geom = "text", base_sizes = c(5,8,5))
pbiplot_vertebrates
ggsave(pbiplot_vertebrates, filename = "CCA_biplot_Vertebrates.jpg", dpi = 600, width = 10, height = 10)
```

# Do analysis for **plants** only
```{r, echo = TRUE}
rl_plants <- rl_all %>% 
	filter(Spine == "Plant")
nrow(rl_plants)

table(rl_plants$Biotope)
table(rl_plants$Biome)

rl_plants <- rl_plants %>% 
	filter(Biotope != "Rheophilous" & Biome != "SemiAquatic")
```

```{r, echo = TRUE}
rl_plants_wide <- rl_plants %>%
	select(TaxonomicGroup, SpeciesnameAsPublished, RLC, Biome, Biotope, Cuddliness, Lifespan, Mobility, NutrientLevel) %>%
	filter(RLC != "DD", RLC != "DD(PT)") %>%
	group_by(RLC, Biome, Biotope, Cuddliness, Lifespan, Mobility, NutrientLevel) %>%
	summarise(nspec = n()) %>%
	spread(key = RLC, value = nspec, fill = 0) %>%
	ungroup() %>%
	filter(complete.cases(.))
```

```{r, include = TRUE}
cca1_plants <- cca(rl_plants_wide %>% select(CR:VU) ~ Biome
									 + Biotope
									 #+ Cuddliness
									 + Lifespan
									 + Mobility
									 + NutrientLevel
									 , data = rl_plants_wide)
summary(cca1_plants)
anova(cca1_plants, by = "terms")
```

```{r, echo = TRUE}
pbiplot_plants <- ggbiplot_vegan(cca1_plants, site_geom = "blank", cn_geom = "text", base_sizes = c(5,8,5))
pbiplot_plants
ggsave(pbiplot_plants, filename = "CCA_biplot_Plants.jpg", dpi = 600, width = 10, height = 10)
```
