---
title: "Darwin Core mapping"
subtitle: "For: Validated red lists of Flanders, Belgium"
author:
- Dimitri Brosens
- Peter Desmet
- Lien Reyserhove
- Dirk Maes
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: true
    toc: true
    toc_depth: 3
    toc_float: true
---

This document describes how we map the validated red list data for Flanders to Darwin Core. The source file for this document can be found [here](https://github.com/inbo/rl-flanders-checklist/blob/master/src/dwc_mapping_validated.Rmd).

# Setup

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Load libraries:

```{r}
# devtools::install_github(c("tazinho/snakecase"))

library(tidyverse)      # To do data science
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(digest)         # To generate hashes
library(rgbif)          # To use GBIF services
library(snakecase)      # To convert case of descriptions
```

# Read source data

Create a data frame `input_data` from the source data:

```{r}
input_data <- read_delim(here("data", "raw", "tblFlandersRedListsAll.tsv"), delim = "\t")
```

Filter on **validated** taxa:

```{r}
input_data %<>% filter(Validated == "Validated")
```

Number of records:

```{r}
input_data %>% nrow()
```

# Process source data

## Tidy data

Clean data somewhat:

```{r}
input_data %<>%
  remove_empty("rows") %>%    # Remove empty rows
  clean_names()               # Have sensible (lowercase) column names
```

## Scientific names

The scientific names contain trailing spaces:

```{r}
input_data %<>% mutate(
  speciesname_as_published = str_trim(speciesname_as_published),
  speciesname_unique = str_trim(speciesname_unique)
)
```

## Taxon ranks

Use the [GBIF nameparser](https://www.gbif.org/tools/name-parser) to retrieve nomenclatural information for the scientific names in the checklist:

```{r parse_names}
parsed_names <- input_data %>%
  distinct(speciesname_as_published) %>%
  pull() %>% # Create vector from dataframe
  parsenames() # An rgbif function
```

The nameparser function also provides information about the rank of the taxon (in `rankmarker`). Here we join this information with our checklist. Cleaning these ranks will done in the Taxon Core mapping:

```{r}
input_data %<>% left_join(
  select(parsed_names, scientificname, rankmarker),
  by = c("speciesname_as_published" = "scientificname"))
```

## Preview data

Show the number of taxa per kingdom and rank:

```{r}
input_data %>%
  group_by(kingdom, rankmarker) %>%
  summarize(
    `taxa` = n()
  ) %>%
  adorn_totals("row")
```

Preview data:

```{r}
input_data %>% head()
```

# Taxon core

## Pre-processing

```{r start_taxon}
taxon <- input_data
```

## Term mapping

Map the data to [Darwin Core Taxon](http://rs.gbif.org/core/dwc_taxon_2015-04-24.xml).

### language

```{r}
taxon %<>% mutate(dwc_language = "en")
```

### license

```{r}
taxon %<>% mutate(dwc_license = "http://creativecommons.org/publicdomain/zero/1.0/")
```

### rightsHolder

```{r}
taxon %<>% mutate(dwc_rightsHolder = "INBO")
```

### accessRights

```{r}
taxon %<>% mutate(dwc_accessRights = "https://www.inbo.be/en/norms-data-use")
```

### datasetID

```{r}
taxon %<>% mutate(dwc_datasetID = "")
```

### institutionCode

```{r}
taxon %<>% mutate(dwc_institutionCode = "INBO")
```

### datasetName

```{r}
taxon %<>% mutate(dwc_datasetName = "Validated Red Lists of Flanders, Belgium")
```

### taxonID

```{r}
taxon %<>% mutate(dwc_taxonID = unique_id)
```

### scientificName

Use the name as originally published on the checklist:

```{r}
taxon %<>% mutate(dwc_scientificName = speciesname_as_published)
```

### kingdom

```{r}
taxon %<>% mutate(dwc_kingdom = kingdom)
```

### phylum

```{r}
taxon %<>% mutate(dwc_phylum = phylum)
```

### class

```{r}
taxon %<>% mutate(dwc_class = class)
```

### order

```{r}
taxon %<>% mutate(dwc_order = order)
```

### family

```{r}
taxon %<>% mutate(dwc_family = family)
```

### genus

```{r}
taxon %<>% mutate(dwc_genus = genus)
```

### taxonRank

Inspect values:

```{r}
taxon %>%
  group_by(rankmarker) %>%
  count()
```

Map values by recoding to the [GBIF rank vocabulary](http://rs.gbif.org/vocabulary/gbif/rank_2015-04-24.xml):

```{r}
taxon %<>% mutate(dwc_taxonRank = recode(rankmarker,
  "sp." = "species",
  "infrasp." = "infraspecificname",
  "subsp." = "subspecies",
  "var." = "variety",
  .default = "",
  .missing = ""
))
```

Inspect mapped values: 

```{r}
taxon %>%
  group_by(rankmarker, dwc_taxonRank) %>%
  count()
```

### vernacularName

```{r}
taxon %<>% mutate(dwc_vernacularName = speciesname_dutch)
```

### nomenclaturalCode

```{r}
taxon %<>% mutate(dwc_nomenclaturalCode = case_when(
  kingdom == "Animalia" ~ "ICZN",
  kingdom == "Plantae" ~ "ICBN"
))
```

## Post-processing

Only keep the Darwin Core columns:

```{r}
taxon %<>% select(starts_with("dwc_"))
```

Drop the `dwc_` prefix:

```{r}
colnames(taxon) <- str_replace(colnames(taxon), "dwc_", "")
```

Sort on `taxonID` (to maintain some consistency between updates of the dataset):

```{r}
taxon %<>% arrange(taxonID)
```

Preview data:

```{r}
taxon %>% head()
```

Save to CSV:

```{r}
write_csv(taxon, here("data", "processed", "validated", "taxon.csv"), na = "")
```

# Distribution extension

## Pre-processing

```{r start_distribution}
distribution <- input_data
```

## Term mapping

Map the data to [Species Distribution](http://rs.gbif.org/extension/gbif/1.0/distribution.xml).

### taxonID

```{r}
distribution %<>% mutate(dwc_taxonID = unique_id)
```

### locationID

```{r}
distribution %<>% mutate(dwc_locationID = "ISO_3166-2:BE-VLG")
```

### locality

```{r}
distribution %<>% mutate(dwc_locality = "Flanders")
```

### countryCode

```{r}
distribution %<>% mutate(dwc_countryCode = "BE")
```

### occurrenceStatus

Set to `absent` for regionally extent species, otherwise `present`:

```{r}
distribution %<>% mutate(dwc_occurrenceStatus = recode(rlc, 
  "RE" = "absent",
  .default = "present",
  .missing = "present"
))
```

### threatStatus

There are two red list category columns:

```{r}
distribution %>%
  group_by(rlc, rlc_as_published) %>%
  count()
```

This will be mapped as follows:

- `rlc` → `threatStatus`: IUCN equivalent of Flemish status and according to [expected vocabulary](http://rs.gbif.org/vocabulary/iucn/threat_status.xml).
- `rlc_as_published` → `occurrenceRemarks`: Flemish status as originally published in red list. Not according to vocabulary, but important include.

```{r}
distribution %<>% mutate(dwc_threatStatus = rlc)
```

### establishmentMeans

```{r}
distribution %<>% mutate(dwc_establishmentMeans = case_when(
  rlc_as_published == "Niet-inheemse broedvogel" ~ "introduced",
  TRUE ~ "native"
))
```

### eventDate

```{r}
distribution %<>% mutate(dwc_eventDate = year_published)
```

### source

```{r}
distribution %<>% mutate(dwc_source = reference)
```

### occurrenceRemarks

```{r}
distribution %<>% mutate(dwc_occurrenceRemarks = rlc_as_published)
```
 
## Post-processing

Only keep the Darwin Core columns:

```{r}
distribution %<>% select(starts_with("dwc_"))
```

Drop the `dwc_` prefix:

```{r}
colnames(distribution) <- str_replace(colnames(distribution), "dwc_", "")
```

Sort on `taxonID`:

```{r}
distribution %<>% arrange(taxonID)
```

Preview data:

```{r}
distribution %>% head()
```

Save to CSV:

```{r}
write_csv(distribution, here("data", "processed", "validated", "distribution.csv"), na = "" )
```

## Description extension

## Pre-processing

```{r start_description}
description_ext <- input_data
```

Gather description columns to rows:

```{r}
description_ext %<>% gather(
  key = type, value = description,
  biome, biotope1, biotope2, lifespan, cuddliness, mobility, spine, nutrient_level,
  na.rm = TRUE
)
```

Rename `biotope1` and `biotope2` to `biotope`:

```{r}
description_ext %<>% mutate(type = recode(type,
  "biotope1" = "biotope",
  "biotope2" = "biotope"
))
```

Inspect values:

```{r}
description_ext %>%
  select(type, description) %>%
  group_by(type, description) %>%
  count()
```

Convert descriptions from `CamelCase` to `lower case`:

```{r}
description_ext %<>% mutate(
  clean_description = str_to_lower(to_sentence_case(description)))
```

Inspect mapped values:

```{r}
description_ext %>%
  group_by(description, clean_description) %>%
  count()
```

## Term mapping

Map the data to [Taxon Description](http://rs.gbif.org/extension/gbif/1.0/description.xml).

### taxonID

```{r}
description_ext %<>% mutate(dwc_taxonID = unique_id)
```

### description

```{r}
description_ext %<>% mutate(dwc_description = clean_description)
```

### type

```{r}
description_ext %<>% mutate(dwc_type = case_when(
  type == "nutrient_level" ~ "nutrient level",
  TRUE ~ type
))
```

### source

```{r}
description_ext %<>% mutate(dwc_source = reference)
```

### language

```{r}
description_ext %<>% mutate(dwc_language = "en")
```

## Post-processing

Only keep the Darwin Core columns:

```{r}
description_ext %<>% select(starts_with("dwc_"))
```

Drop the `dwc_` prefix:

```{r}
colnames(description_ext) <- str_replace(colnames(description_ext), "dwc_", "")
```

Sort on `taxonID`:

```{r}
description_ext %<>% arrange(taxonID)
```

Preview data:

```{r}
description_ext %>% head()
```

Save to CSV:

```{r}
write_csv(description_ext, here("data", "processed", "validated", "description.csv"), na = "" )
```
