---
title: "Darwin Core mapping"
subtitle: "For: Unvalidated red lists of Flanders, Belgium"
author:
- Dimitri Brosens
- Peter Desmet
- Lien Reyserhove
- Dirk Maes
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: true
    toc: true
    toc_depth: 3
    toc_float: true
---

This document describes how we map the unvalidated red list data for Flanders to Darwin Core. The source file for this document can be found [here](https://github.com/inbo/rl-flanders-checklist/blob/master/src/dwc_mapping_unvalidated.Rmd).

# Setup

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Load libraries:

```{r}
# devtools::install_github(c("tazinho/snakecase"))

library(tidyverse)      # To do data science
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(digest)         # To generate hashes
library(rgbif)          # To use GBIF services
library(snakecase)      # To convert case of descriptions
```

# Read source data

Create a data frame `input_data` from the source data:

```{r}
input_data <- read_delim(here("data", "raw", "unvalidated", "NonValidatedRedListsSource.csv"), delim = ";")
```

Preview data:

```{r}
input_data %>% head()
```

# Process source data

## Tidy data

Clean data somewhat:

```{r}
input_data %<>%
  remove_empty("rows") %>%    # Remove empty rows
  clean_names()               # Have sensible (lowercase) column names
```

## Get GBIF Backbone Taxonomy information

Get GBIF Backbone Taxonomy information to map taxon rank and higher classification. This will overwrite any existing column of the same name (e.g. `genus`):

```{r gbif_match_names}
input_data %<>% gbif_species_name_match(
  name_col = "speciesname_as_published", 
  gbif_terms = c(
    "usageKey", "kingdom", "phylum", "class", "order", "family", "genus", 
    "scientificName", "rank", "status", "matchType", "confidence"
  )
)
```

## Preview data

Show the number of taxa per kingdom and rank:

```{r}
input_data %>%
  group_by(kingdom, rank) %>%
  summarize(
    `taxa` = n()
  ) %>%
  adorn_totals("row")
```

Preview data:

```{r}
input_data %>% head()
```

# Taxon core

## Pre-processing

```{r start_taxon}
taxon <- input_data
```

## Term mapping

Map the data to [Darwin Core Taxon](http://rs.gbif.org/core/dwc_taxon_2015-04-24.xml).

### language

```{r}
taxon %<>% mutate(dwc_language = "en")
```

### license

```{r}
taxon %<>% mutate(dwc_license = "http://creativecommons.org/publicdomain/zero/1.0/")
```

### rightsHolder

```{r}
taxon %<>% mutate(dwc_rightsHolder = "INBO")
```

### accessRights

```{r}
taxon %<>% mutate(dwc_accessRights = "https://www.inbo.be/en/norms-data-use")
```

### datasetID

```{r}
taxon %<>% mutate(dwc_datasetID = "")
```

### institutionCode

```{r}
taxon %<>% mutate(dwc_institutionCode = "INBO")
```

### datasetName

```{r}
taxon %<>% mutate(dwc_datasetName = "Unvalidated Red Lists of Flanders, Belgium")
```

### taxonID

```{r}
taxon %<>% mutate(dwc_taxonID = unique_id)
```

### scientificName

```{r}
taxon %<>% mutate(dwc_scientificName = speciesname_as_published)
```

### kingdom

```{r}
taxon %<>% mutate(dwc_kingdom = kingdom)
```

### phylum

```{r}
taxon %<>% mutate(dwc_phylum = phylum)
```

### class

```{r}
taxon %<>% mutate(dwc_class = class)
```

### order

```{r}
taxon %<>% mutate(dwc_order = order)
```

### family

```{r}
taxon %<>% mutate(dwc_family = family)
```

### genus

```{r}
taxon %<>% mutate(dwc_genus = genus)
```

### taxonRank

Inspect values:

```{r}
taxon %>%
  group_by(rank) %>%
  count()
```

```{r}
taxon %<>% mutate(dwc_taxonRank = recode(rank,
  "GENUS" = "genus",
  "SPECIES" = "species",
  "SUBSPECIES" = "subspecies",
  "VARIETY" = "variety",
  .missing = ""
))
taxon %<>% mutate(dwc_taxonRank = "species")
```

### vernacularName

```{r}
taxon %<>% mutate(dwc_vernacularName = speciesname_dutch)
```

### nomenclaturalCode

```{r}
taxon %<>% mutate(dwc_nomenclaturalCode = case_when(
  kingdom == "Animalia" ~ "ICZN",
  kingdom == "Plantae" ~ "ICBN"
))
```

## Post-processing

Only keep the Darwin Core columns:

```{r}
taxon %<>% select(starts_with("dwc_"))
```

Drop the `dwc_` prefix:

```{r}
colnames(taxon) <- str_replace(colnames(taxon), "dwc_", "")
```

Sort on `taxonID` (to maintain some consistency between updates of the dataset):

```{r}
taxon %<>% arrange(taxonID)
```

Preview data:

```{r}
taxon %>% head()
```

Save to CSV:

```{r}
write_csv(taxon, here("data", "processed", "unvalidated", "taxon.csv"), na = "")
```

# Distribution extension

## Pre-processing

```{r start_distribution}
distribution <- input_data
```

## Term mapping

Map the data to [Species Distribution](http://rs.gbif.org/extension/gbif/1.0/distribution.xml).

### taxonID

```{r}
distribution %<>% mutate(dwc_taxonID = unique_id)
```

### locationID

```{r}
distribution %<>% mutate(dwc_locationID = "ISO_3166-2:BE-VLG")
```

### locality

```{r}
distribution %<>% mutate(dwc_locality = "Flanders")
```

### countryCode

```{r}
distribution %<>% mutate(dwc_countryCode = "BE")
```

### occurrenceStatus

Set to `absent` for regionally extent species, otherwise `present`:

```{r eval = FALSE}
distribution %<>% mutate(dwc_occurrenceStatus = recode(rlc, 
  "RE" = "absent",
  .default = "present",
  .missing = "present"
))
distribution %<>% mutate(dwc_occurrenceStatus = "present")
```

### threatStatus

There are two red list values, which we will map as follows:

- `rlc_as_published`: Flemish status as officially published. Will be mapped _as is_ to `occurrenceRemarks` so this information is included.
- `rlc`: IUCN equivalent of Flemish status. Will be mapped to `threatStatus` (which expects this controlled vocabulary).

```{r eval = FALSE}
distribution %>%
  group_by(rlc_as_published, rlc) %>%
  count()
```

```{r eval = FALSE}
distribution %<>% mutate(dwc_threatStatus = rlc)
```

```{r}
distribution %<>% mutate(dwc_threatStatus = recode(rlc_as_published
      , "Geografisch beperkt" = "Geographically Limited (NT)"
      , "Endangered (EN), maar niet gekend in welke mate" = "Endangered (EN), but not known to what extent"
      , "Near Threatened (NT) (vrij zeldzaam)"="Near Threatened (NT) (quite rare)"
      , "Near Threatened (NT) (zeer zeldzaam)"="Near Threatened (NT) (very rare)"
      , "Near Threatened (NT) (zeldzaam)"="Near Threatened (NT) (rare)"
      , "Niet bedreigd" = "Least Concern (LC)"
      , "Least concern"="Least Concern (LC)"
      , "Sterk bedreigd"="Critically Endangered (EN)"
      , "Vulnerable" = "Vulnerable (VU)"
      , "Verdwenen"= "Disappeared (RE)"
      , "Waarschijnlijk bedreigd"="Probably Endangered (NT)"
      , "Vermoedelijk bedreigd"="Presumably Threatened (DD/PT)"
      , "Vatbaar voor bedreiging"="Vulnerable to threat (NT)"
      , "Critically endangered"="Critically Endangered (CR)"
      , "Critically Endangered"="Critically Endangered (CR)"
      , "Endangered"= "Endangered (EN)"
      , "Regionally extinct"="Regionally extinct (RE)"
      , "Disappeared"="Disapeared (RE)"
      , "Near threatened"="Near Threatened (NT)"
      , "Probably Endangered"="Probably Endangered (DD/PT)"
      , "Geographically Limited"="Geographically Limited (NT)" 
      , "Bedreigd"="Endangered (EN)" 
      , "Kwetsbaar"= "Vulnerable (VU)"
      , "Ernstig bedreigd"="Critically Endangered (CR)"
      , "Momenteel niet bedreigd"="Least Concern (LC)"
      , "Momenteel niet in gevaar"="Least Concern (LC)"
      , "Met uitsterven bedreigd"="Critically Endangered (CR)"
      , "Bijna in gevaar"="Near Threatened (NT)"
      , "Met uitsterven bedreigd"="Critically Endangered (CR)"
      , "Zeldzaam"="Near Threatened (NT)"
      , "Uitgestorven"="Regionally Extinct (REX)"
      , "Achteruitgaand"="Near Threatened (NT)"
      , "Onvoldoende gekend"="Data Deficient (DD)"
      , "Onvoldoende data "= "Data Deficient (DD)"
      , "Regionaal uitgestorven"="Regionally Extinct (EX)"
      , "maar mate waarin ongekend"="uncertain rate"
      , "in Vlaanderen"="in Flanders"
))
```

### establishmentMeans

```{r}
distribution %<>% mutate(dwc_establishmentMeans = "native")
```

### eventDate

```{r}
distribution %<>% mutate(dwc_eventDate = year_published)
```

### source

```{r}
distribution %<>% mutate(dwc_source = reference)
```

### occurrenceRemarks

```{r}
distribution %<>% mutate(dwc_occurrenceRemarks = rlc_as_published)
```

### extra

```{r}
distribution %<>% mutate(
  dwc_redListCriteria = rlc_iucn,
  dwc_criteria = criteria,
  dwc_redListCriteriaEurope = rlc_europe
)
```

## Post-processing

Only keep the Darwin Core columns:

```{r}
distribution %<>% select(starts_with("dwc_"))
```

Drop the `dwc_` prefix:

```{r}
colnames(distribution) <- str_replace(colnames(distribution), "dwc_", "")
```

Sort on `taxonID`:

```{r}
distribution %<>% arrange(taxonID)
```

Preview data:

```{r}
distribution %>% head()
```

Order columns to compare in git:

```{r}
distribution %<>% select(
  taxonID,redListCriteria,redListCriteriaEurope,criteria,eventDate,source,locality,occurrenceStatus,countryCode,locationID,threatStatus,establishmentMeans
)
```

Save to CSV:

```{r}
write_csv(distribution, here("data", "processed", "unvalidated", "distribution.csv"), na = "" )
```
